### 2013-09-09 thesis/src/chapters/chapter-2.tex
Vertical scaling

\subsection{Horizontal scaling}
As outlined in previous section, horizontal scaling is all about adding supplementary nodes to a system. As it is common to cloud computing, nodes are represented as virtual machines and this assumption is used in further discussion. As a consequence, adding server comes down to cloning new virtual machine from a template and possibly installing additional software or reconfiguring it later. While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack \cite{OpenStack}, OpenNebula \cite{OpenNebula}, CloudStack \cite{CloudStack} or Eucalyptus \cite{Eucalyptus} to name a few) and is similar in manner, the underlying virtualization mechanism influences how fast provisioning is done. Table \ref{tab:hypervisors-provisioning} describes in details different capabilities of commonly used virtualization technologies.

@@

\subsubsection*{Load-balancing algorithms}
Generally, there are two types of load-balancers: hardware and software based. Due to the dynamic nature of system under consideration, we focus only on the latter as it offers a greater deal of flexibility. Among the most common algorithms we can distinguish:
\begin{itemize}
 \item \textit{round-robin scheduling} -
 \item \textit{first in first out} -
 \item \textit{shortest remaining time} -
\end{itemize}

Situation gets further complicated when considering real-world web application that sends user information using cookies, what imposes requirement on load-balancer for session stickiness \cite{StBaMa11}. Table \ref{tab:load-balancers} summaries available load-balancing mechanism including their utilization in IaaS products.

\begin{table}[!htbp]
\begin{tabularx}{\textwidth}{l | X | X}
 &  & \\
\hline 
HAProxy &  & \\
\hline
Apache modproxy &  & \\
\hline
... &  & \\
\end{tabularx}
\caption{Comparision of load balancers}
\label{tab:load-balancers}
\end{table}

\subsubsection*{Load-balancing scalability}
Although, it may seem that balancing workloads eliminates problem of a single point of failure (SPOF) among different servers, it is in fact shifted to load-balancing layer. In other words, load-balancer becomes a new SPOF. Therefore, in cases where high availability is required, multi-tiered load balancing architecture should be considered. This, however, seems not to be a case among IaaS or PaaS providers - none of them unequivocally specifies whether their provide redundancy at load-balancer level \textbf{TODO check :)}

>>>>

\subsection{Horizontal scaling}
As outlined in previous section, horizontal scaling is all about adding supplementary nodes to a system. As it is common to cloud computing, nodes are represented as virtual machines and this assumption is used in further discussion. As a consequence, adding server comes down to cloning new virtual machine from a template and possibly installing additional software or reconfiguring it later. While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack \cite{OpenStack}, OpenNebula \cite{OpenNebula}, CloudStack \cite{CloudStack} or Eucalyptus \cite{Eucalyptus} to name a few) and is similar in manner, the underlying virtualization mechanism determines how fast provisioning is done. Table \ref{tab:hypervisors-provisioning} describes in details different capabilities of commonly used virtualization technologies.

@@

\subsubsection{Load-balancing algorithms}
Generally, there are two types of load-balancers: hardware and software based. Due to the dynamic nature of system under consideration, we focus only on the latter as it offers a greater deal of flexibility. Among the most common algorithms we can distinguish:
\begin{itemize}
 \item \textit{round-robin scheduling} - 
 \item \textit{least connection} - 
 \item \textit{source routing} - 
 \item \textit{URI hashing} - 
\end{itemize}

Situation gets further complicated when considering real-world web application that sends user information using cookies, what imposes requirement on load-balancer for session stickiness \cite{StBaMa11}. Below are presented one of the most common load-balancers, including their key performance features.

\subsubsection*{HAProxy}

HAProxy \cite{HAProxy} is a load-balancer developed by. Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears (CITE NEEDED). Its key features include:
\begin{itemize}
	\item a single-process, event-driven model reduces the cost of context switch and the memory usage
	\item O(1) event checker
	\item Single-buffering without copying data between reads and writes
	\item Zero-copy forwarding
	\item Optimized HTTP header analysis : headers are parsed an interpreted on the fly
\end{itemize}

Requests are dispatched accordingly to chosen algorithm:
\begin{itemize}
 \item \textit{round-robin scheduling}
 \item \textit{least connection}
 \item \textit{source routing}
 \item \textit{URI hashing}
\end{itemize}

\subsubsection*{Apache HTTP with modproxy}

\subsubsection*{Apache HTTP with mod_jk}

\subsubsection{Load-balancing scalability}
Although, it may seem that balancing workloads eliminates problem of a single point of failure (SPOF) among different servers, it is in fact shifted to load-balancing layer. In other words, load-balancer becomes a new SPOF. Therefore, in cases where high availability is required, multi-tiered load balancing architecture should be considered. This, however, seems not to be a case among IaaS or PaaS providers - none of them unequivocally specifies whether their provide redundancy at load-balancer level \textbf{TODO check :)}
